### 2015-11-04 paper.tex
a wimpy conclusion

\documentclass{sig-alternate}
\usepackage{color}

\newcommand{\red}[1]{\textcolor{red}{#1}}

%\title{Adaptation Writ Large:\\ An Essay in Memory of John H. Holland}

\title{Adaptative Computation:\\ An Essay in Memory of John H. Holland}

\numberofauthors{2}
\author 
{\alignauthor
 Stephanie Forrest\\
 \affaddr{University of New Mexico}\\
 \affaddr{Santa Fe Institute}\\
 \email{forrest@cs.unm.edu}
 \alignauthor
Melanie Mitchell \\
 \affaddr{Portland State University}\\
 \affaddr{Santa Fe Institute}\\
 \email{mm@pdx.edu}
}

\begin{document}
\maketitle

% \begin{abstract}
% \input{abstract}
% \end{abstract}

\section{Introduction}

Professor John H. Holland passed away recently in Ann Arbor, MI, where
he had been on the University of Michigan faculty for over 50 years.  John, as he was known
universally to his colleagues and students, leaves behind a long
legacy of intellectual achievements.

As a descendant of the cybernetics era, Holland was strongly influenced
by the work of von Neumann, Wiener, Ashby, and Turing, all of whom
viewed computation as a broad, interdisciplinary enterprise.  Thus,
Holland became an early proponent of interdisciplinary approaches to
computer science and an active evangelist of what is now called
\emph{computational thinking}, reaching out enthusiastically to
psychologists, economists, physicists, linguists, philosophers, and
pretty much anyone he came in contact with.  As a result, even though
he received what was arguably the world's first computer science
Ph.D. in 1959, his contributions are sometimes better known outside of
computer science than within.

Holland is best known for his invention of \emph{genetic
  algorithms} (GAs), a family of search and optimization methods inspired
by biological evolution.  Since their invention in the 1960s, GAs have
inspired many related methods and led to the thriving subfield of
\emph{evolutionary computation}, with widespread scientific and
commercial applications.  While the mechanisms of GAs are well-known,
%to much of the computer science community, 
the ideas that motivated Holland are less widely appreciated---to develop a general
%Holland's motivation 
%offshoot of Holland's much broader motivation--
theory of adaptation in complex systems.  
%This motivation was the
%driving force in all of Holland's research.

In this short essay, we consider this larger framework, sketching the
recurring themes that were central to Holland's theory of
adaptive systems: (1) discovery and 
%evolutionary 
dynamics of building blocks; (2) internal models and prediction; (3) 
exploratory modeling; and (4) complex adaptive systems.  
%We discuss the role these
%themes have played in computer science, and highlight especially the
%ideas that we think remain relevant to today's research agendas.  [MM
%says: Are we going to do this?]

%\section{Evolutionary Dynamics of ``Building Blocks''}  
\section{Discovery and Dynamics of Building Blocks}

Holland's interest in a general theory of adaptation grew out of
his early work developing computer models of Hebbian learning
\cite{Rochester1956} and 
his reading of 
Ronald Fisher's classic work that
integrated genetics with Darwinian selection \cite{Fisher1930}.
%As Holland further read extensively in 
Over time, he recognized that adaptation was central to 
evolutionary biology,
economics, psychology, game theory, control theory, and many other fields.
%he came to recognize that
%adaptation was the most central concept in all these fields.  That is,
Each field concerns populations of agents that must continually
obtain information from uncertain changing environments and use it to
improve their performance and enhance survival.
%with respect to those environments.
Moreover, Holland recognized that adaptation must be a continual
open-ended process due to the
environment's \emph{perpetual novelty}.
%SF thinks Holland did not invent this term (as Holland termed it), 
Thus, adaptive systems never achieve a state of equilibrium or a final
optimum configuration.  This emphasis on open-ended, non-equilibrium
dynamics was in stark contrast with the mainstream approach (at the
time) in all these fields---the belief that solving for stable
equilibrium dynamics was the scientific goal.  Holland's contrary view
was that a system in stable equilibrium is essentially \emph{dead}.

% Text that STEPH found in our NIPS paper on the IGA
% First, at least theoretically the GA is fast because of implicit
% parallelism (Holland, 1975/1992): each string in the population is an
% instance of many different schemas, and if the population is large
% enough and is initially chosen at random, a large number of different
% schemas-many more than the number of strings in the population-are
% being sampled in parallel.  This should result in a quick search for
% short, low-order schemas that confer high fitness.  Second,
% fitness-proportionate reproduction under the GA should conserve
% instances of such schemas.  Third, a high crossover rate should
% quickly combine instances oflow-order schemas on different strings to
% create instances of longer schemas that confer even higher fitness.

Underlying Holland's theory of adaptation are the following core ideas:

\begin{itemize}
\item{\bf Populations and implicit parallelism:} Adaptation occurs in
  populations of individuals evolving (or learning) over time, in
  which statistics can be leveraged to direct population dynamics.
  Evolution can be thought of as a search for populations that are highly fit with respect to their environment.  
  A population can be thought of as a parallel sampling of many
  individuals (from the space of possible individuals).  Moreover,
  that same population can be thought of as implicitly sampling a much
  space of \emph{traits} that comprise those individuals.
  Holland termed this implicit large-scale sampling of traits ``implicit
  parallelism.''    Evolutionary dynamics biases these samples over time towards
  high-fitness regions of the search space.

%The population can be
%  thought of as a set of samples in a much larger search space, where
%  the processes of evolution
%  population can be thought of as a set of sa in which statistics 
%  can be leveraged to direct population dynamics.  (More on this below.)

\item{\bf Building blocks and recombination:} 
% Melanie rewrote this some to take out references to ``solutions''
  In a population undergoing adaptation,
  individuals can be decomposed into \emph{building blocks}---sets of
  traits that are the evolutionary ``atoms'' of an individual's
  fitness or performance. (As an example from biology, Holland gives
  the \emph{Krebs cycle}, a core cellular metabolic
  pathway.)   Successful individuals
  are discovered in stages, first by finding useful
  building blocks through random sampling, and over
  time recombining such building blocks to create higher-fitness individuals.

%  For example, when searching the space of
%  all $n$-bit binary strings, almost any random sample would contain
%  examples of strings whose first bit is set to 1, and also many
%  samples of strings whose second bit is set to 1, but there would be
%  many fewer samples that have both of the first two bits set to
%  1.  More compelling examples include the \emph{Krebs cycle}, a core
%  cellular metabolic pathway,

  To succeed, such a search requires first that
  useful building blocks be preserved once discovered, and second that
  there be an efficient mechanism for combining promising building
  blocks into higher-order building blocks.  Holland proposed using an
  analog of Darwinian selection for the first and the idea of crossing
  over in genetics for the second.  Holland referred to the building
  blocks as \emph{schemas} and defined them as hyperplanes in the search space.

%In a population undergoing adaptation,
%  individuals can be decomposed into \emph{building blocks}---sets of
% traits that are the evolutionary ``atoms'' of an individual's
%  fitness or performance.

\item{\bf Exploitation Versus Exploration:} Successful adaptation
  requires maintaining a balance between \emph{exploitation}, in which
 successful building blocks propagate in a population, and \emph{
    exploration}, in which existing building blocks are recombined or
  mutated in new ways.  Inspired by Bellman \cite{Bellman1961} and
others, Holland formalized this exploitation-versus-exploration tradeoff
via an idealized ``two-armed bandit'' problem.  Given a slot machine
with two arms, each of which has an unknown payoff probability, how
should you allocate $N$ trials (pulls) between the arms so as to
maximize your total payoff?   In \cite{Holland1973,Holland1975}
Holland 
%derived
%an equation for such a strategy and 
argued that the optimal strategy
allocates trials to the observed best arm at a slightly higher rate than
exponential.
% and extended the result to multi-armed bandits.
%A system that focuses too much on exploitation
%  risks never finding better individuals or not being able to adapt to
%  changing environments.  Conversely, too much focus on exploration
%  means that the system won't leverage the successful building blocks
%  that have already been discovered.  %There is some optimal balance
%  between these two modes of processing.
Next, he showed that 
populations undergoing adaptation resemble multi-armed bandits, where
each building block is an 
%undergoing adaptation constitute the 
arms in a multi-armed bandit problem.
%Each building block can be said to have a probability of ``payoff''
%(i.e., contribution to a given individual's fitness).
Thus, in the process of assigning a fitness to each individual in a
population, adaptive evolution can be seen as implicitly sampling the
many building blocks making up that collection of individuals.  The
average fitness over many individuals containing a particular building
block gives an estimate of that building block's payoff probability.

\end{itemize}

Several of Holland's early papers (e.g.,
\cite{Holland1962,Holland1973}) and his influential
1975 book \emph{Adaptation in Natural and Artificial Systems}
\cite{Holland1975} developed a general, formal setting in which these ideas
could be expressed mathematically.

It was this formalization
% and resulting theorems that led to the
that led to the invention of genetic algorithms, which featured stochastic
population-based search, as well as
%recombination as a critical
crossover between individuals as a critical
operation that allowed successful building blocks to be recombined and
tested in new contexts.  What made GAs unique
among other evolution-inspired algorithms at the time were the
mathematical foundations described above, the emphasis on populations
and recombination as central mechanisms,
and a focus 
%(at least in Holland's mind) [MM SAYS: THIS SOUNDS TOO NEGATIVE]
on continual adaptation to
non-stationary environments rather than optimization to static
environments.

The framework Holland developed in \cite{Holland1975} was more general
than the genetic algorithm; its aim was an interdisciplinary theory of
adaptation, one that would inform biology, say, as much as computer
science \cite{Christiansen1998}.  The later, successful application of
genetic algorithms to real-world optimization and learning tasks was,
for Holland, just icing on the cake.

%\section{Internal Models and Perpetual Novelty}
\section{Internal Models and Prediction}

%One system of special interest to Holland was the human mind, and in
%particular, its capacity to learn via induction.

Internal models are central to Holland's theory of adaptive
systems.  He posits that all adaptive systems create and use internal
models to prosper in the face of ``perpetual novelty'' in continually changing environments.   Models can be tacit and learned
  through evolutionary time, as in the case of bacteria swimming up a
  chemical gradient, or explicit and learned over a single lifespan,
  as in the case of cognitive systems that incorporate experience into
  internal representations through learning.  In
  Holland's view, the key activity of an adaptive agent involves
  building and refining these data-driven models
  of the environment.
%using them to make predictions and take actions,
%  and occasionally receiving positive or negative feedbacks
%  (intermittent rewards).

In his second book, \emph{Induction} \cite{Holland1989}, Holland and
his co-authors tackled the question of how cognitive agents can
possibly learn and profit from internal models by combining
environmental inputs and rewards with stored knowledge.  In their
framework, a model defines a set of equivalence relations over
environmental states, together with a set of transition rules, which
are learned over time based on environmental rewards (or punishments).
Models that form valid homomorphisms with the environment allow the
system to make accurate predictions.  In Holland's conception, the
equivalence classes are initially very general, say ``moving object''
and ``stationary object.'' Through experience and
learning, these can be specialized into more useful and precise subclasses, say
``insect'' and ``nest.''  Over time, the adaptive system builds up a
default hierarchy of rules covering general cases and refinements for
specific classes.

% For example, a general rule might
% state that, ``If X is a bird, then X can fly'', whereas a more specific
% one might state, ``If X is a bird with small wings and a large body,
% then X cannot fly'' \cite{Holland1989}.  When faced with a bird in
% the environment, both rules might compete to post their conclusion,
% but the more specific one would be favored if both of its conditions are met.

% Maybe this should be moved to section on Complexity
Although the idea of default hierarchies was prevalent in 
knowledge representation systems of the era, Holland made two key
contributions.  The first was his emphasis on homomorphisms as a formal way to
evaluate model validity, an idea that dates back to Ross Ashby's
\emph{An Introduction to Cybernetics} \cite{Ashby1956}. Holland's student
Bernard Ziegler developed this idea into a formal theory of computer modeling
and simulation \cite{Ziegler1976}.  
% %Today, even though computational modeling is a major
% %application area of computing, essential to most natural sciences and
% %engineering disciplines
Even today, these early homomorphic theories of modeling
stand as the most elegant approach we know of to characterize when
a model is consistent with the environment and how an intelligent
agent, human or artificial, can update the model to better reflect
reality.

Holland's second key contribution was describing a computational
mechanism, the \emph{learning classifier
  system}~\cite{Holland1977,Holland1986}, to illustrate how a cognitive system could
iteratively build up a detailed and hierarchical model of its
environment to enhance survival.  The key learning elements of this method, the
bucket-brigade algorithm combined with a genetic algorithm, presaged many of the ideas in modern reinforcement
learning.

%including unsupervised learning, non-Markovian learning, AND
%ONE OTHER?.

% While genetic algorithms are nowadays most often associated with models of
% biological evolution, in Holland's view the processes of adaptation he
% was studying manifested themselves in many different complex systems.

% In the 1970s and 80s Holland formulated and extensively explored a
% particular model of decision-making, action, and inductive learning,
% called the \emph{classifier system} (e.g.,)
% .  In a classifier system, a population
% of ``if-then'' rules interacts with an environment, and over time is
% able to learn to improve its performance via both \emph{
%   credit-assignment} and \emph{rule-discovery} algorithms.

% Like genetic algorithms, classifier systems can be viewed both as 
% models of adaptation and as artificial-intelligence (AI) methods.  At
% the time Holland was developing classifier systems, the field of AI
% was focused on expert systems, which typically did not learn on their
% own, a fundamental deficiency in Holland's view: ``[Expert] systems
% are brittle in the sense that they respond appropriately only in
% narrow domains, requiring substantial human intervention to compensate
% for even slight shifts in domain'' \cite{Holland1986}. In
% contrast, Holland proposed that ``\emph{induction} is the basic, and
% perhaps only, way of making large advances in this direction.''

Holland's inspiration for classifier systems came from several
different disciplines, including Hebbian
learning, artificial intelligence, evolutionary biology, economics,
psychology, control theory, and other fields (e.g.,
\cite{Bellman1961,Samuel1959}).  Knowledge representation in the form of a
population of ``if-then'' rules seemed like a good choice, not only
because of its popularity in AI at the time but also from Holland's
early work on modeling Hebbian cell assemblies: ``In Hebb's view, a
cell assembly makes a simple statement: If such and such an event
occurs, then I will fire for a while at a high
rate.'' \cite{Waldrop1993} (p. 182).  
The if-then rules, when activated, compete to post their results on a
shared ``message list,'' modeling the system's short-term memory
(again inspired by Hebb's work and AI blackboard systems of the day).
Unlike the AI systems, however, new rules were generated automatically
%in a trial-and-error fashion
using a genetic algorithm.

Successful rules were strengthened over time if their predictions led
to positive rewards from the environment (and weakened otherwise)
through a credit-assignment method called the \emph{bucket-brigade}
algorithm, in which rules gaining rewards from the environment or from
other rules transferred some of their gains to those earlier-firing
``stage-setting'' rules that set up the conditions for the eventual
reward. Holland credited Arthur Samuel's pioneering work on machine
learning applied to checkers \cite{Samuel1959} as a key inspiration for these ideas.

%, e.g., finding a target, or avoiding a predator,

%Finally, new rules for the population are discovered by a genetic
%algorithm, using the strength of rules as a fitness function.
%Classifier systems are fairly complex architectures, with learning
%occurring at multiple time scales.  
Although Holland proposed classifier systems as 
an executable theory of inductive processes in cognition, other researchers took it further, 
%other researchers took the
%computational system more seriously
applying it to areas as diverse
as poker-playing \cite{Smith1980}, control of gas pipeline
transmission \cite{Goldberg1983}, and modeling the stock market
\cite{Palmer1994}.  (See Ref.~\cite{Booker1989} for more details about
practical applications of classifier systems.)  Today, other
reinforcement learning methods are more popular for real-world
decision and control problems, but classifier systems can perhaps be
thought of as an essential stage-setting method that enabled the
development of later approaches.

% \footnote{Of course similar
%   credit-assignment methods have been used extensively in
%   reinforcement learning, neural networks, and other areas of modern
%   machine learning, but Holland was among the first to propose such
%   methods.}

Holland was primarily interested in how the two
learning mechanisms (discovery of new rules and apportioning credit to
existing rules) could work together to create 
%subsets of rules
%that accurately model the environment, and thus are able to ``fire''
%at appropriate times.  
%Holland envisioned that, with a complex environment and enough time
%for learning, a classifier system might learn an intricate default
useful default hierarchies of rules.  He emphasized that the
competition inherent in the learning and action mechanisms would allow
the system to adapt to a continually changing environment without
losing what it had learned in the past.  Holland put it this way:
``Competition among rules provides the system with a graceful way of
handling perpetual novelty.  When a system has strong rules that
respond to a particular situation, that is the equivalent of saying
that it has certain well-validated hypotheses.... New rules do not
interfere with the system's action in well-practiced situations but
wait gracefully in the wings as hypotheses about what to do under
novel circumstances.'' \cite{Holland1992}.

%Offspring rules, which begin life weaker than do their parents, can
%win the competition and influence the system's behavior only when
%there are not strong rules whose conditions are satisfied---in other
%words when the system does not know what to do.  If their actions
%help, they survive; if not they are soon replaced.  
%Thus,

\section{Exploratory Modeling}

Given that Holland believed that the ability to learn and manipulate
internal models was essential for any adaptive system, it is no
surprise that he viewed modeling as essential for scientific inquiry.

%Today, we use computational models in several distinct ways---as
%tools for analyzing data in statistical models; as methods for
%discovering new knowledge; and as a means to understand how complex
%systems works.  [Steph, I am not clear on the distinction between the
%second and third things here, so I combined them.  Change back if you
%want.]

Today, we use computational models both for \emph{prediction}---by
analyzing data via statistical models---and for \emph{understanding}
how systems work---by probing the effects of hypothesized underlying
mechanisms.  This latter use of models was dear to Holland's heart.
In his view, the key to science was understanding the mechanisms that
cause a system to behave in a certain way, an aspiration that goes
well beyond data fitting methods, which typically focus only on
the aggregate behavior of a system.

For example, a purely statistical model that describes the boom and bust
pattern of the stock market
%cannot
does not address the underlying mechanisms
that lead to these cycles, through the collective actions of myriad
individual buy/sell decisions.  
%Similarly
In contrast, the genetic algorithms for which Holland is so famous
are exploratory models of \emph{mechanism}: they provide a simple
computational framework in which to explore the dynamics of Darwinian
evolution and whether the basic mechanisms of variation, differential
reproduction, and heredity are sufficient to account for the richness
of our natural world.

Holland was interested in models that explored basic principles and
mechanisms, even if they did not make specific or detailed
predictions.  Such models can show generically how certain behaviors
could be produced.  Holland pioneered a style of modeling that has
come to be known as ``individual-based'' or ``agent-based,'' in which
every component of a system is represented explicitly---e.g., every
trader in a stock market system or every cell in an immune system
model---and has a dynamic internal state. In such models, each agent has
its own behavior rules, which it can modify over time through learning.  In
order to capture the behavior of systems under spatial
constraints, these models are often defined over spatial structures,
such as networks or simple grids.

A given agent-based model encodes a theory about the mechanisms that
are relevant for producing the behavior of interest.  
%Similar to expert systems [MM says: ?]
%Such models are especially useful for studying systems
%that do not have analytic mathematical descriptions.  [MM took out to shorten this paragraph]
%Agent-based models can facilitate interdisciplinary collaborations because the
%underlying rules can be easily communicated.  
%[SF: THINKS this next text doesn't make sense until after complex
%systems seciton.]
The exploratory models
championed by Holland were idealized and not intended to provide detailed, domain-specific
predictions.  Instead they were meant to explore possible general
mechanisms of complex systems and thus provide insights that might
lead to more specific, detailed models.  Such idealized models are
akin to what Dennett has called ``intuition pumps''
\cite{Dennett1984}.

The emphasis on exploratory models to build intuitions was an
important theme of Holland's work, and he often quoted Eddington's
remark on the occasion of the first experimental test of Einstein's
theory of relativity: ``The contemplation in natural science of a wider
domain than the actual leads to a far better understanding of the
actual'' \cite{Eddington1927}.

It should be noted that Holland's view of modeling is by no means
typical.  For example, a textbook on computational modeling offers
 the following definition: ``Modeling is the application
of methods to analyze complex, real-world problems in order to make
predictions about what might happen with various actions''
\cite{Shiflet2014}.   This perspective rules out the
kind of exploratory modeling that Holland was most interested in.

% Some researchers dispute that models make any kind of scientific
% contribution: ``Models are metaphors that explain the world we don't
% understand in terms of worlds we do.  They are merely analogies,
% provide partial insight, stand on someone else's feet.  Theories stand
% on their own feet, and rely on no analogies.''  [Emanuel Derman, 2012].
% [MM SAYS:  I COULD NOT FIND REFERENCE FOR THIS QUOTE OUTSIDE SFI VIDEO.]
% [MM SAYS:  Possibly delete this paragraph.]

% [NOW WE NEED A STRONG FINISHING SENTENCE TO RESCUE JHH STYLE
%   MODELING.]

%\section{Common Properties of Complex Adaptive Systems}

\section{Complex Adaptive Systems}

%\section{Complex Adaptive Systems}

Holland was interested in a broad array of adaptive systems---immune
systems, ecologies, financial markets, cities, and the brain---systems
that are \emph{complex}.  In the 1980s, he teamed up with a small
group of scientists, primarily physicists with a sprinkling of
economists and biologists, to discuss what properties this wide swath of systems
have in common.  The discussions helped define the intellectual mission
of the Santa Fe Institute (SFI), the first institution dedicated to
developing a science of complexity, and the many others that followed.
% [add in how many complexity
%  institutes there are worldwide and which others John played a hand
%  in. MM says: Do we really need that?]  
Holland brought to these discussions his lifelong study of
adaptation and a reminder that serious theories about complexity would
need to look deeper than phenomenological descriptions but also
account for the `how' and `why' of these systems.
%describing phenomena like chaos, power laws, and ONE MORE, and acoount for the

As the discussions matured, a consensus developed about the basic
elements of complex systems, which are: (1) 
%complex systems are 
composed of many
components with nonlinear interactions; (2) 
%such systems
%are 
characterized by complex \emph{emergent} behavior, exhibiting
higher-order patterns; (3) 
%such systems 
operating at multiple (and often
% interacting? 
nested) spatial and temporal scales, with some behavior being
conserved across all scales and other behaviors changing at different
scales, and (4) 
%these systems exhibit 
adaptative, with 
behavioral rules continually adjusting through evolution and learning.  Although this
is far from a formal definition of complex systems, most people
working in the field today are interested in systems that have these
properties.

%To illustrate his ideas about the ubiquity of adaptation, Holland and
%co-workers developed two exploratory models, the SFI Artificial Stock
%Market and ECHO.

In the early 1990s, Holland teamed up with other Santa Fe Institute researchers,
including several economists, to tackle the mismatch between predictions
of rational expectations theory (the then dominant theory in economics) 
and empirically observed stock market behaviors. 
%that deviated from the predicted equilibrium solutions.  
In brief, most
economic theory of the day assumed that all participants in an economy
or financial market are 100\% rational and act to maximize their
individual gain.  In real life, however, actors in economies
and markets are rarely wholly rational, and financial markets often deviate
from rationality, for example with speculative bubbles and crashes.

The SFI Artificial Stock Market project \cite{Arthur1997,Palmer1994}
developed an exploratory model in which rational
traders were replaced by \emph{adaptive} traders---those who learn to
forecast stock prices over time.  The model tested for the
emergence of different trading strategies, including
\emph{fundamental}, \emph{technical}, or \emph{uninformed}
strategies. The simulated market with adaptive trading agents was run
many times, and the dynamics of price and trading volumes were
compared to observed patterns in real markets.  Holland and his
collaborators found that the model's dynamics replicated several
features of real-life markets.

Although the SFI Stock Market model was highly simplified, it was very
influential and led to many follow-on projects.  It 
demonstrated clearly the essential role that adaptation plays in complex
systems, and it illustrated how Holland's theories of continual learning
in response to intermittent feedbacks from the environment could be
integrated into domain-specific settings.

ECHO \cite{Holland1999,Hraber1997} was an even more ambitious model
that Holland and collaborators developed during the 1990s. 
% ECHO's
% scope included the evolution of interacting populations of individual
% agents, which through competition and learning discover symbiotic
% triads such as the famous ant-fly-caterpillar interaction, the Wiksell
% triangle of trading relationships in economics, and the immune
% system's learned ability to distinguish `self' from `other.'
ECHO formalized Holland's idealization of complex adaptive systems into
a runnable computational system where agents evolved external markers
(called tags) and internal preferences, and then used them to acquire
resources, form higher
level aggregate structures (such trading relationships, symbiotic groups,
trophic cascades, interdependent organizations, etc.).  ECHO agents
sometimes discovered mimicry to deceive competitors, and over time,
the model developed increasingly complex structures and behaviors, at
least up to a point.  ECHO reproduced several well-known patterns 
observed in nature, for example, over time the population of ECHO
evolved agents exhibited diversity of agent types, diversity with a 
rank-frequency distribution that
closely paralleled the well-known Preston curve in ecology.  In
layman's terms, ECHO evolved populations where 
%quantitative statement of the adage that
``most species are rare.''  
The broad scope of the model, together with its ability to produce
easily identifiable patterns, was appealing to immunologists, economists, and evolutionary biologists alike.   Many
of the insights behind this project are described in Holland's third book \emph{Hidden Order} \cite{Holland1995}.

% but it
%was the economists who took it one step further in the SFI stock
%market.

%Blake LeBaron (economics);
% W. Brian Arthur (economics); 
% John Holland (psychology/EE/CS,
% and father of GAs);
% Richard Palmer (physics);
% Paul Taylor (computer science

%Many modeling issues not satisfactorily resolved by the SF-ASM model have been 
% taken up in later research (see Ref.[4]).

Holland's later books, \emph{Emergence} \cite{Holland2000},
\emph{Signals and Boundaries} \cite{Holland2012}, and
\emph{Complexity: A Very Short Introduction} \cite{Holland2014} show how the
theories of adaptation that Holland developed during the earlier part
of his career fit into the larger landscape of complex systems
research.  Holland's focus on understanding the mechanisms by which complex patterns emerge and change,
rather than simply characterizing the patterns themselves (e.g.,
describing chaotic attractors or power laws), reflected his
determination to get to the heart of complex adaptive systems.
This determination represents the best of science.  Holland's
willingness to tackle the hardest questions, develop his own
formalisms, and use mathematics productively sets a high bar to whch we
all should aspire.

%\section{Relevance to Modern CS}
%[MM says:  Do we need this section?  Or can we incorporate these points within the previous sections?]

\section{Conclusion}

Holland was unusual in his ability to absorb the essence of other
disciplines, articulate grand overarching
theories, and them back them up with computational mechanisms and
mathematics.  Unlike most researchers he moved seamlessly among these
three modes of thinking, developing models that were years ahead of their time.  A
close reading of his work reveals the antecedents of many ideas prevalent in
machine learning today, e.g., reinforcement, learning in non-Markovian
environments, active learning, etc.   His seminal genetic algorithm
spawned the field of evolutionary computation, and his insights
and wisdom helped define what are now referred to as the
\emph{sciences of complexity}.   With such a broad scope, it is easy
to pick at the details.

His many books and articles have influenced 
scientists around the work and across many disciplines.  Closer to home, he introduced several
generations of students at the University of Michigan to 
computation in natural systems, an idea that even today
remains somewhat controversial, despite successes in biomimicry
for robotics, abstractions of phermone trails in ant colonies for
optimization methods, the use of mechanisms from immunology to
improve computer security, and recent work on technological
evolution.

Behind the ideas, of course, is the man himself.   Everyone who knew
John personally will remember the gleam in his eye when he
encountered a new idea; his willingness to talk to anyone, no matter
how famous or how unfamous; and his incredible generosity and patience.
%His insights aboutwere deeper and more general than what often passes for work in biomimicry, e.g., for robots.
Holland's personality and humanity were somehow inextricably tangled
up with his intellectual contributions, and although his ideas will serve as a beacon
for research in intelligent and complex systems
for many years to come, the world of science is poorer without John
Holland on the scene.

% \begin{itemize}
% \item Evolutionary computation
% \item Q Learning
% \item Backprop
% \item    From Turing nomination: "Holland's machine learning system known as the
%    Learning Classifier System (LCS), developed in the early 1980s,
%    incorporated a reinforcement learning algorithm known as the bucket
%    brigade for non-Markovian environments, anticipating by nearly a
%    decade non-Markovian learning algorithms."
% \item Exploitation versus exploration -- relevance to reinforcement
%   learning, other parts of machine learning and optimization.
% \item Active learning.  (Two-armed bandit problem.) 
% \item On-line learning.  
% \end{itemize}

\bibliographystyle{plain}

\bibliography{paper}

\end{document}